\fred{A Unified Treatment of Tractabilty of Approximation Problems}\\ \peter{[PETER: In view of the papers of Henryk and Michael, maybe we should think of a better title?]}

### 2023-06-19 GeneralizedTractability_230607.tex
Updates from Overleaf

\documentclass[11pt,a4paper]{article}
\usepackage{latexsym,amsfonts,amsmath,graphics,mathtools,amssymb,amsthm,booktabs}
\usepackage{epsfig}
\usepackage[notref,notcite]{showkeys}
%\usepackage[active]{srcltx}
\usepackage{enumitem}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[normalem]{ulem}
\usepackage{algorithmic}

@@

@@

\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}

\title{Generalized Tractability}
\author{\fred{Fred J. Hickernell}, \peter{Peter Kritzer}, \kachi{Onyekachi Osisiogu}, and \\ Henryk Wo\'zniakwoski}
\date{\today}

@@

Let $\{\cf_d\}_{d \in \naturals}$ and $\{\cg_d\}_{d \in \naturals}$ be sequences of Hilbert spaces, and $\{\SOL_d : \cf_d \to \cg_d\}_{d \in \naturals}$ be a sequence of linear solution operators  with adjoints $\SOL_d^*$ such that $\SOL_d^*\SOL_d : \cf_d \to \cf_d$ has eigenvalues and orthonormal eigenvectors
\begin{equation} \label{eq:eigenvalues}
 \lambda_{1,d}^2 \ge \lambda_{2,d}^2 \ge \cdots, \qquad u_{1,d}, u_{2,d}, \ldots, \qquad d \in \naturals.
\end{equation}
This means that the non-negative $\lambda_{i,d}$ are the singular values of the operator $\SOL_d$. We remark that in the literature on Information-Based Complexity the singular values of $\SOL_d$ are often denoted by $\sqrt{\lambda_{i,d}}$ instead of $\lambda_{i,d}$. However, since we assume non-negativity of the $\lambda_{i,d}$, the notation used here is equivalent. 
\peter{[PETER: I am wondering whether we should go back to the more common notation once the manuscript is finalized (not right away). It would make the paper easier to read for the IBC community.]}\fred{Fred: If we do, we introduce a lot of $\sqrt{\lambda}$ or $\varepsilon^{-2}$, which I find harder to translate in my brain.  But I  will submit to the will of the majority.}
The non-decreasing ordering of the eigenvalues for each $d$ facilitates the proof of conditions for tractability, but our conditions do not depend on the ordering.

We make the technical assumption that $\SOL_d$ is a compact operator. This implies (see, e.g., \cite{NW08}) that $\lambda_{i,d}$ converges to zero as $i$ tends to infinity for every $d$, which in turn implies that the problem is solvable by suitable algorithms. To avoid trivial cases, we assume that there are an infinite number of positive eigenvalues for every $d$.

@@

The question of tractability is one of determining how fast the complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  We want to bound the complexity in terms of a simple function of $\varepsilon^{-1}$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding.

In this article, we investigate such a function that we refer to as the generalized tractability function, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty),
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we we define our approximation problem to be tractable if 
$\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant, $C_{\vp}$, depending only on the parameter $\vp$.  In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase mildly with increasing $\varepsilon^{-1}$ and/or $d$, however most of our theorems do not require such an assumption.

@@

Certain common notions of tractability  covered by the general situation described above are provided in Table \ref{tab:commonT}.  Here, $\vp = (p,q)$ in some cases and $\vp$ is a scalar in other cases. A general reference
for results on algebraic tractability is \cite{NW08}, for exponential tractability we refer to \cite{HKW19} and \cite{KW19}.

@@

Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},d',\vp)$ are non-decreasing in $i$ for any $d'\in\naturals$. In particular, for  any $N\in\naturals$ with $n\ge N$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{N,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},d',\vp)} \le \cdots \le \frac{1}{T(\lambda_{N, d}^{-1},d',\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp) }
    \le &\frac{1}{n-N+1} \sum_{i=N}^n  \frac{1}{T(\lambda_{i, d}^{-1},d',\vp) }
    \le \frac{1}{n-N+1} \sum_{i=N}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)}.
\end{align*}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n - N +1 \ge T(\varepsilon^{-1},d',\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp)} \le
   \frac{1}{n-N+1} \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp) } \le \frac{1}{T(\varepsilon^{-1},d',\vp)} \\ \forall d', n, N \in \naturals.
\end{multline}
We note that the left side of the implication implies $n \ge N$ automatically since $n$ is an integer  $n - N +1$ is assumed to be positive.

Moreover, given that $T$ is increasing in its arguments, we have the following equivalent expression for the complexity via \eqref{eq:comp_def}:
\begin{align} %\label{eq:comp_equiv}
	\nonumber
	\comp(\varepsilon, d) & = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} \\
	\nonumber
	& = \min \biggl\{n \in \natzero : \frac{1}{T(\lambda_{n+1,d}^{-1},d',\vp)}\le \frac{1}{T(\varepsilon^{-1},d',\vp) }\biggr\} \qquad  \forall d'\in \naturals\\
	\nonumber
	& \le  \min \biggl\{n \in \natzero : n \ge N - 1 + T(\varepsilon^{-1},d',\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \biggr\} \;  \forall  d', N \in \naturals, \quad \text{by \eqref{eq:implicate_a}} \\ 
	& \le  T(\varepsilon^{-1},d',\vp) \Biggl [\frac{N}{T(\varepsilon^{-1},d',\vp)}  +  \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \Biggr] \quad \forall  d', N \in \naturals. \label{eq:compUP}
\end{align}

Now we take this upper bound further, specializing to the case of $d'=1$ and $N=L_{\vp}$:
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},1, \vp) \Biggl [ \frac{L_{\vp}}{T(\varepsilon^{-1},1, \vp)}  + \underbrace{\sum_{i=L_{\vp}}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}}_{\le S_{\vp} <\infty\ \text{by \eqref{eq:strong_tractiff3}} }
        \Biggr] \\
       & \le T(\varepsilon^{-1},1,\vp) \underbrace{\left[\frac{L_{\vp}}{T(0,1, \vp)} + S_{\vp} \right]}_{=:C_{\vp}}
       \qquad \text{by \eqref{eq:Tnondecreasing}}\\
       & =  C_{\vp}T(\varepsilon^{-1},1,\vp).
%       & \le \underbrace{2 C_{1,\vp}}_{\text{since } T(\varepsilon^{-1},1,\vp) \ge 1} T(\varepsilon^{-1},1,\vp)
%       = \underbrace{C_{\vp}}_{:=2 C_{1,\vp}}T(\varepsilon^{-1},1,\vp).
\end{align*}
This means that we have strong $T$-tractability via \eqref{eq:strong_tract_def}, and verifies the sufficiency of \eqref{eq:strong_tractiff3}.

@@

\noindent This concludes the proof of  Theorem \ref{thm_main_strong_tract2}.

@@

\begin{proof}
    \textbf{Sufficient Condition:}\\
Suppose that \eqref{eq:tractiff4} holds for some $\vp \ge 0$. 
By the argument in the proof of the sufficient condition for Theorem \ref{thm_main_strong_tract2}, it follows from \eqref{eq:compUP} with $d'=d$ and $N = \lceil L_{\vp}\, T(0,d,\vp)\rceil$ that 
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},d,\vp) \Biggl [ 
       \frac{\lceil L_{\vp}\, T(0,d,\vp)\rceil}{T(\varepsilon^{-1},d,\vp)}  + 
       \underbrace{\sum_{i=\lceil L_{\vp}\, T(0,d,\vp)\rceil}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}}
       \Biggr ] 
        \quad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)+1}{T(\varepsilon^{-1},d, \vp)} +  S_{\vp} \right]
        \qquad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)}{T(\varepsilon^{-1},d, \vp)}
       + \frac{1}{T(0,1, \vp)} + S_{\vp}\right]
       \qquad \text{by \eqref{eq:Tnondecreasing}}   \\
       & \le T(\varepsilon^{-1},d,\vp) 
       \underbrace{\left [  L_{\vp} + \frac{1}{T(0,1, \vp)} + S_{\vp} \right]}_{=:C_{\vp}}
       \qquad \text{by \eqref{eq:Tnondecreasing}}  
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}.

@@

Thus,  $\cp_{\textup{opt}}= \widetilde{\cp}_{\textup{opt}}$.  This concludes the proof of Theorem \ref{thm_main_tract2}.
\end{proof}

\section{Examples} \label{sec:examples}

In this section, we illustrate Theorems \ref{thm_main_strong_tract2} and \ref{thm_main_tract2} for the types of tractability given in the table in Section \ref{sec:not}.

\begin{example}[Strong algebraic polynomial tractability]

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-p}\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]  where in this case $\vp$ is a scalar.
Then Theorem \ref{thm_main_strong_tract2} yields the  equivalent condition for strong algebraic polynomial tractability:
\[
 \sup_{d\in\naturals} \sum_{i=L_p}^\infty (\min\{1,\lambda_{i,d}^p\}) < \infty \text{ for some } p \ge 0, L_p \in \naturals.
\]
Furthermore, the optimal $p$ is the infimum of all $p$ for which the above condition holds.
This essentially recovers the result in \cite[Theorem 5.1]{NW08}.
\end{example}

\begin{example}[Algebraic polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-p}\}\, d^q
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where in this case $\vp = (p,q)$.
Then Theorem \ref{thm_main_tract2} yields the  equivalent condition for algebraic polynomial tractability:
\[
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty (\min\{1,\lambda_{i,d}^p\})  < \infty \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
Furthermore, the set of optimal $(p,q)$ is a subset of
\begin{multline*}
\{(p^*,q^*) : \text{the above condition holds for all }(p,q) \in (p^*,\infty) \times (q^*,\infty) \\
\text{but not for any }(p,q) \in [0,p^*/2) \times [0,q^*/2)\}.
\end{multline*}
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NW08} \fred{same theorem as above??, what about optimality in \cite[Theorem 5.1]{NW08}}.
\end{example}

@@

As adefined in \cite{GW11}, we say that a problem is \emph{quasi-polynomially tractable (QPT)} if there exist two constants $C$, $p > 0$, such that
\begin{equation}\label{expqpt}
    \comp(\varepsilon,d) \leq C\exp\{p(1+\log(\varepsilon^{-1}))(1+\log(d))\}
\end{equation}
for all $\varepsilon \in (0,1]$ and $d\in\mathbb{N}$.

@@

\begin{theorem}\label{thm:subhT}
	Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions}.  Moreover, let $\lim_{d \to \infty} T(\varepsilon^{-1},d,\vp) = \infty$ for all $\varepsilon > 0$, $\vp > \vzero$.  The problem is sub-$h$-$T$ tractable if and only if
	\[
	S_c = S_{c,\vp} :=\sup_{d\in\naturals}  \, \sum_{i=1}^\infty \frac1{h(c\, T(\lambda_{i,d}^{-1},d,\vp))} <\infty \qquad \forall c > 0.
	\]
\end{theorem}
\begin{proof}
The idea of this proof follows \cite{WW17}.
We suppress the $\vp$ dependence in the proof, because it insignificant.

@@

Suppose that the domain of interest $(\varepsilon^{-1},d)$ is not all of $(0,\infty) \times \naturals$ but some subset
\begin{equation}
    \Omega  := \{(\varepsilon^{-1},d) : \varepsilon \in (E(d),\infty), \ d \in \naturals \}, \qquad \text{where }E : \naturals \to [0,\infty).
\end{equation}
We then expect that the equivalent conditions for (strong) tractability will be similar in form, but weaker than, the conditions in Theorems  \ref{thm_main_strong_tract2} and \ref{thm_main_tract2}.

@@

Define $M: \naturals \to \natzero \cup \{\infty\}$
\begin{equation} \label{eq:Mddef}
    M(d) := \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le E(d)\}, & E(d) > 0, \\
        \infty, & E(d) = 0.
    \end{cases}
\end{equation}
Also, for each $d \in \naturals$ define the interval 
\begin{equation} \label{eq:hIdef}
\hI_{d}:=\begin{cases} 
\emptyset, & M(d) = 0, \\
\{1,\ldots, M(d)\}, & 0 < M(d) < \infty, \\
\naturals, & M(d) = \infty.
\end{cases}
\end{equation}
From the definition of $M(d)$ it follows that $(\lambda_{i,d}^{-1},d) \in \Omega  \iff i \in \hI_d$, and
\begin{equation} \label{eq:comprestr}
  \comp(\varepsilon,d) \begin{cases}
= 0, & M(d) = 0, \\
    \in \hI_d, & M(d) > 0, 
\end{cases}
\qquad \forall (\varepsilon^{-1},d) \in \Omega.
\end{equation}

The equivalent conditions for (strong) $T$-tractability on a restricted domain in Theorems \ref{thm_main_rest_strong_tract} and \ref{thm_main_rest_tract} mimic the conditions given in Theorems \ref{thm_main_strong_tract2} and \ref{thm_main_tract2}, except that the upper limits on the sums now correspond to $M(d)$ rather than $\infty$.  The proofs are a bit more delicate, but similar arguments are used.

>>>>

\documentclass[sort&compress]{elsarticle}
\usepackage{latexsym,amsfonts,amsmath,graphics,mathtools,amssymb,booktabs,xspace}
\usepackage{epsfig}
\usepackage[notref,notcite]{showkeys}
%\usepackage[active]{srcltx}
\usepackage{fdsymbol}
\usepackage{enumitem}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[normalem]{ulem}
\usepackage{algorithmic}
\usepackage{hyperref}
\usepackage{cleveref}

@@

\newtheorem{theorem}{Theorem}
\newdefinition{definition}{Definition}
\newdefinition{example}{Example}
\newproof{proof}{Proof}

@@

\title{\fred{A Unified Treatment of Tractabilty of Approximation Problems}\\ \peter{[PETER: In view of the papers of Henryk and Michael, maybe we should think of a better title?]}}

\author[1]{Fred J. Hickernell}
\ead{hickernell@iit.edu}
\author[2]{Peter Kritzer}
\author[1]{Onyekachi Osisiogu}
\date{\today}
\affiliation[1]{organization={Department of Applied Mathematics, Illinois Institute of Technology},
addressline={10 West 32${}^{\text{nd}}$ Street, Room 220},
postcode={60616},
city={Chicago, IL},
country={United States}}

\begin{abstract}
    A large literature specifies conditions under which the information complexity for a sequence of numerical problems defined for dimensions $1, 2, \ldots$ grows at a moderate rate, i.e., the sequence of problems is \emph{tractable}.  Here we focus on the situation where the space of available information consists of all linear functionals and the problems are defined as operator mappings between Hilbert spaces.  We unify the proofs of known tractability results and generalize a number of existing results.  These generalizations are expressed as five theorems that provide equivalent conditions for (strong) tractability in terms of sums of functions of the singular values.  
\end{abstract}

@@

The information complexity of a problem is the number of function data required to solve the problem within the desired error tolerance.  A problem is tractable if the information complexity does not grow too large as the dimension of the problem increases or the error threshold decreases.  There are a number of tractability results for numerical problems beginning with \cite{?}, spawning numerous articles, and filling a series of three impressive volumes \cite{NovWoz08a,NovWoz10a,NovWoz12a}.  The purpose of this article is to unify the many of these tractability results for approximation problems where all linear functional data is available.

Let $\{\cf_d\}_{d \in \naturals}$ and $\{\cg_d\}_{d \in \naturals}$ be sequences of Hilbert spaces, and $\{\SOL_d : \cf_d \to \cg_d\}_{d \in \naturals}$ be a sequence of linear solution operators  with adjoints $\SOL_d^*$ such that $\SOL_d^*\SOL_d : \cf_d \to \cf_d$ has eigenvalues and orthonormal eigenvectors
\begin{equation} \label{eq:eigenvalues}
 \lambda_{1,d}^2 \ge \lambda_{2,d}^2 \ge \cdots, \qquad u_{1,d}, u_{2,d}, \ldots, \qquad d \in \naturals.
\end{equation}
This means that the non-negative $\lambda_{i,d}$ are the singular values of the operator $\SOL_d$. We remark that in the literature on Information-Based Complexity the singular values of $\SOL_d$ are often denoted by $\sqrt{\lambda_{i,d}}$ instead of $\lambda_{i,d}$. However, since we assume non-negativity of the $\lambda_{i,d}$, the notation used here is equivalent.

%The non-decreasing ordering of the eigenvalues for each $d$ facilitates the proof of conditions for tractability, but our conditions do not depend on the ordering \peter{[PETER: is that really the case?]}.

We make the technical assumption that $\SOL_d$ is a compact operator. This implies (see, e.g., \cite{NovWoz08a}) that $\lambda_{i,d}$ converges to zero as $i$ tends to infinity for every $d$, which in turn implies that the problem is solvable by suitable algorithms. To avoid trivial cases, we assume that there are an infinite number of positive eigenvalues for every $d$.

@@

The question of tractability is one of determining how fast the complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  We want to bound the complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding.

In this article, we investigate a \emph{generalized tractability function}, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we we define our approximation problem to be tractable if 
$\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant, $C_{\vp}$, depending only on the parameter $\vp$.  In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase mildly with increasing $\varepsilon^{-1}$ and/or $d$, however most of our theorems do not require such an assumption.

\Cref{thm_main_strong_tract2,thm_main_tract2,thm_main_rest_strong_tract,thm_main_rest_tract} all take the form
\begin{multline}
  \exists C_{\vp} \text{ with }
    \comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},\clubsuit,\vp) \quad \forall (\varepsilon^{-1},d) \in \Omega \\
    \iff   \exists L_{\vp} \text{ with }
    \sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,\clubsuit,\vp) \rceil}^{\spadesuit(d)} \frac{1}{T(\lambda^{-1}_{i,d},\clubsuit,\vp)}< \infty, \\
     \clubsuit \in \{1,d\}, \ \varheartsuit(d) = \inf \{\varepsilon : (\varepsilon^{-1},d) \in \Omega \}, \ 
     \spadesuit(d) = \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le \varheartsuit(d) \}, &   \varheartsuit(d)< 0, \\
        \infty, & \varheartsuit(d) = 0.
        \end{cases}
\end{multline}
\Cref{thm:subhT} is of similar, but not equivalent, form.

@@

Certain common notions of tractability  covered by the general situation described above are provided in Table \ref{tab:commonT}.  Here, $\vp = (p,q)$ in some cases and $\vp$ is a scalar in other cases. A general reference
for results on algebraic tractability is \cite{NovWoz08a}, for exponential tractability we refer to \cite{HicKriWoz19a} and \cite{KriWoz19a}.

@@

Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},\clubsuit,\vp)$ are non-decreasing in $i$ for any $\clubsuit\in\naturals$. In particular, for  any $N\in\naturals$ with $n\ge N$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{N,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\clubsuit,\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},\clubsuit,\vp)} \le \cdots \le \frac{1}{T(\lambda_{N, d}^{-1},\clubsuit,\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\clubsuit,\vp) }
    \le &\frac{1}{n-N+1} \sum_{i=N}^n  \frac{1}{T(\lambda_{i, d}^{-1},\clubsuit,\vp) }
    \le \frac{1}{n-N+1} \sum_{i=N}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},\clubsuit,\vp)}.
\end{align*}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n - N +1 \ge T(\varepsilon^{-1},\clubsuit,\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\clubsuit,\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},\clubsuit,\vp)} \le
   \frac{1}{n-N+1} \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\clubsuit,\vp) } \le \frac{1}{T(\varepsilon^{-1},\clubsuit,\vp)} \\ \forall \clubsuit, n, N \in \naturals.
\end{multline}
We note that the left side of the implication implies $n \ge N$ automatically since $n$ is an integer  $n - N +1$ is assumed to be positive.

Moreover, given that $T$ is increasing in its arguments, we have the following equivalent expression for the complexity via \eqref{eq:comp_def}:
\begin{align} %\label{eq:comp_equiv}
	\nonumber
	\comp(\varepsilon, d) & = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} \\
	\nonumber
	& = \min \biggl\{n \in \natzero : \frac{1}{T(\lambda_{n+1,d}^{-1},\clubsuit,\vp)}\le \frac{1}{T(\varepsilon^{-1},\clubsuit,\vp) }\biggr\} \qquad  \forall \clubsuit\in \naturals\\
	\nonumber
	& \le  \min \biggl\{n \in \natzero : n \ge N - 1 + T(\varepsilon^{-1},\clubsuit,\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\clubsuit,\vp)} \biggr\} \;  \forall  \clubsuit, N \in \naturals, \quad \text{by \eqref{eq:implicate_a}} \\ 
	& \le  T(\varepsilon^{-1},\clubsuit,\vp) \Biggl [\frac{N}{T(\varepsilon^{-1},\clubsuit,\vp)}  +  \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\clubsuit,\vp)} \Biggr] \quad \forall  \clubsuit, N \in \naturals. \label{eq:compUP}
\end{align}

Now we take this upper bound further, specializing to the case of $\clubsuit=1$ and $N=L_{\vp}$:
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},1, \vp) \Biggl [ \frac{L_{\vp}}{T(\varepsilon^{-1},1, \vp)}  + \underbrace{\sum_{i=L_{\vp}}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}}_{\le S_{\vp} <\infty\ \text{by \eqref{eq:strong_tractiff3}} }
        \Biggr] \\
       & \le T(\varepsilon^{-1},1,\vp) \underbrace{\left[\frac{L_{\vp}}{T(0,1, \vp)} + S_{\vp} \right]}_{=:C_{\vp}}
       \qquad \text{by \eqref{eq:Tnondecreasing}}\\
       & =  C_{\vp}T(\varepsilon^{-1},1,\vp).
%       & \le \underbrace{2 C_{1,\vp}}_{\text{since } T(\varepsilon^{-1},1,\vp) \ge 1} T(\varepsilon^{-1},1,\vp)
%       = \underbrace{C_{\vp}}_{:=2 C_{1,\vp}}T(\varepsilon^{-1},1,\vp).
\end{align*}
This means that we have strong $T$-tractability via \eqref{eq:strong_tract_def}, and verifies the sufficiency of \eqref{eq:strong_tractiff3}.

@@

\noindent This concludes the proof of  Theorem \ref{thm_main_strong_tract2}. \qed

@@

\begin{proof}
    \textbf{Sufficient Condition:}\\
Suppose that \eqref{eq:tractiff4} holds for some $\vp \ge 0$. 
By the argument in the proof of the sufficient condition for Theorem \ref{thm_main_strong_tract2}, it follows from \eqref{eq:compUP} with $\clubsuit=d$ and $N = \lceil L_{\vp}\, T(0,d,\vp)\rceil$ that 
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},d,\vp) \Biggl [ 
       \frac{\lceil L_{\vp}\, T(0,d,\vp)\rceil}{T(\varepsilon^{-1},d,\vp)}  + 
       \underbrace{\sum_{i=\lceil L_{\vp}\, T(0,d,\vp)\rceil}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}}
       \Biggr ] 
        \quad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)+1}{T(\varepsilon^{-1},d, \vp)} +  S_{\vp} \right]
        \qquad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)}{T(\varepsilon^{-1},d, \vp)}
       + \frac{1}{T(0,1, \vp)} + S_{\vp}\right]
       \qquad \text{by \eqref{eq:Tnondecreasing}}   \\
       & \le T(\varepsilon^{-1},d,\vp) 
       \underbrace{\left [  L_{\vp} + \frac{1}{T(0,1, \vp)} + S_{\vp} \right]}_{=:C_{\vp}}
       \qquad \text{by \eqref{eq:Tnondecreasing}}  
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}.

@@

Thus,  $\cp_{\textup{opt}}= \widetilde{\cp}_{\textup{opt}}$.  This concludes the proof of Theorem \ref{thm_main_tract2}.  \qed
\end{proof}

\section{Examples} \label{sec:examples}
%In this section, we illustrate Theorems \ref{thm_main_strong_tract2} and \ref{thm_main_tract2} for the types of tractability given in the table in Section \ref{sec:not}.
In this section, we present six examples related to Theorem \ref{thm_main_strong_tract2} and Theorem \ref{thm_main_tract2} introduced in Section \ref{sec:spt}. We will continue to present examples in later sections.

In Section 2.2 we considered strong tractability. 
\begin{example}[Strong algebraic polynomial tractability]

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-p}\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]  where in this case $\vp$ is a scalar.
Then Theorem \ref{thm_main_strong_tract2} yields the  equivalent condition for strong algebraic polynomial tractability:
\[
 \sup_{d\in\naturals} \sum_{i=L_p}^\infty (\min\{1,\lambda_{i,d}^p\}) < \infty \text{ for some } p \ge 0, L_p \in \naturals.
\]
Furthermore, the optimal $p$ is the infimum of all $p$ for which the above condition holds.
This essentially recovers the result in \cite[Theorem 5.1]{NovWoz08a}.
\end{example}

\begin{example}[Algebraic polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-p}\}\, d^q
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where in this case $\vp = (p,q)$.
Then Theorem \ref{thm_main_tract2} yields the  equivalent condition for algebraic polynomial tractability:
\[
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty (\min\{1,\lambda_{i,d}^p\})  < \infty \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
Furthermore, the set of optimal $(p,q)$ is a subset of
\begin{multline*}
\{(p^*,q^*) : \text{the above condition holds for all }(p,q) \in (p^*,\infty) \times (q^*,\infty) \\
\text{but not for any }(p,q) \in [0,p^*/2) \times [0,q^*/2)\}.
\end{multline*}
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NovWoz08a} %\fred{same theorem as above??, what about optimality in \cite[Theorem 5.1]{NovWoz08a}}.
\end{example}

@@

As defined in \cite{GneWoz11a}, we say that a problem is \emph{quasi-polynomially tractable (QPT)} if there exist two constants $C$, $p > 0$, such that
\begin{equation}\label{expqpt}
    \comp(\varepsilon,d) \leq C\exp\{p(1+\log(\varepsilon^{-1}))(1+\log(d))\}
\end{equation}
for all $\varepsilon \in (0,1]$ and $d\in\mathbb{N}$.

@@

\begin{theorem}\label{thm:subhT}
	Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions}.  Moreover, let $\lim_{d \to \infty} T(\varepsilon^{-1},d,\vp) = \infty$ for all $\varepsilon > 0$, $\vp > \vzero$.  The problem is sub-$h$-$T$ tractable if and only if
	\[
	S_c = S_{c,\vp} :=\sup_{d\in\naturals}  \, \sum_{i=1}^\infty \frac1{h(c\, T(\lambda_{i,d}^{-1},d,\vp))} <\infty \qquad \forall c > 0.
	\]
\end{theorem}
\begin{proof}
The idea of this proof follows \cite{WerWoz17a}.
We suppress the $\vp$ dependence in the proof, because it insignificant.

@@

Suppose that the domain of interest $(\varepsilon^{-1},d)$ is not all of $(0,\infty) \times \naturals$ but some subset
\begin{equation}
    \Omega  := \{(\varepsilon^{-1},d) : \varepsilon \in (\varheartsuit(d),\infty), \ d \in \naturals \}, \qquad \text{where }E : \naturals \to [0,\infty).
\end{equation}
We then expect that the equivalent conditions for (strong) tractability will be similar in form, but weaker than, the conditions in Theorems  \ref{thm_main_strong_tract2} and \ref{thm_main_tract2}.

@@

Define $M: \naturals \to \natzero \cup \{\infty\}$
\begin{equation} \label{eq:Mddef}
    \spadesuit(d) := \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le \varheartsuit(d)\}, & \varheartsuit(d) > 0, \\
        \infty, & \varheartsuit(d) = 0.
    \end{cases}
\end{equation}
Also, for each $d \in \naturals$ define the interval 
\begin{equation} \label{eq:hIdef}
\hI_{d}:=\begin{cases} 
\emptyset, & \spadesuit(d) = 0, \\
\{1,\ldots, \spadesuit(d)\}, & 0 < \spadesuit(d) < \infty, \\
\naturals, & \spadesuit(d) = \infty.
\end{cases}
\end{equation}
From the definition of $\spadesuit(d)$ it follows that $(\lambda_{i,d}^{-1},d) \in \Omega  \iff i \in \hI_d$, and
\begin{equation} \label{eq:comprestr}
  \comp(\varepsilon,d) \begin{cases}
= 0, & \spadesuit(d) = 0, \\
    \in \hI_d, & \spadesuit(d) > 0, 
\end{cases}
\qquad \forall (\varepsilon^{-1},d) \in \Omega.
\end{equation}

The equivalent conditions for (strong) $T$-tractability on a restricted domain in Theorems \ref{thm_main_rest_strong_tract} and \ref{thm_main_rest_tract} mimic the conditions given in Theorems \ref{thm_main_strong_tract2} and \ref{thm_main_tract2}, except that the upper limits on the sums now correspond to $\spadesuit(d)$ rather than $\infty$.  The proofs are a bit more delicate, but similar arguments are used.
